---
output: github_document
---

<!-- Note: edit the .Rmd file not the .md file -->

## 1.1

### 1.1a

1. Get into groups of 2 and, in those groups:
1. Create a new RStudio project
1. Create an R script
1. Identify and interact with each of the 4 panels in RStudio
1. Get help on the plot function with `?plot`
1. Create a plot using the `plot()` function
1. Find and install a new package on a topic of your choice with `Tools > Install Packages` (requires internet)
1. Attach the package using `library()`
1. Find and install a new package with `install.packages()`
1. In your source panel write code that creates vector objects `x` and `y` and plots them with `plot(x, y)` to create something that looks like this:
<!-- (is it reproducible?) -->

```{r}
# hint: create an object called x with the following code:
x = c(1, 2, 3, 6)
```

```{r, echo=FALSE}
y = x^2
plot(x, y)
```

- Bonus: find out exactly what R version you are using (tip: use a search engine!)

- Bonus: use R to find out how many minutes you've been alive for. Feel free to use an invented age. Tip: try using both 'base' `as.POSIXct()` and 'tidyverse' `ymd_hm()` functions - you may also need to search online for this.
    
```{r, eval=FALSE, echo=FALSE}
date_of_birth1 = as.POSIXct("1985-12-20 12:00")
date_of_birth2 = ISOdatetime(
  year = 1985,
  month = 12,
  day = 20,
  hour = 12,
  min = 0,
  sec = 0
)
date_of_birth3 = lubridate::ymd_hm("1985-12-20 12:00")
current_time = Sys.time()
diff_days = current_time - date_of_birth1
diff_days * 24 * 60
difftime(current_time, date_of_birth1, units = "m")
```


```{r, echo=FALSE, eval=FALSE}
devtools::install_github("ITSLeeds/stats19")
stats19::dl_stats19_2017_ac()
ac17 = stats19::read_stats19_2005_2014_ac(data_dir = "dftRoadSafetyData_Accidents_2017", filename = "Acc.csv")
ac17 = stats19::format_stats19_2016_ac(ac17)
ac17
View(ac17)
ac_wy = ac17[ac17$Police_Force == "West Yorkshire", ]
ac_wy = dplyr::filter(ac17, Police_)
library(tidyverse)
ac_wy = ac17 %>% 
  filter(Police_Force == "West Yorkshire")
plot(ac_wy$Longitude, ac_wy$Latitude)
```

### 1.1b

1. What class is each of these objects:
    ```{r}
    x = 1:6
    y = x^2
    z = y + 0.1
    ```
    ```{r, echo=FALSE, eval=FALSE}
    class(x)
    typeof(x)
    is.vector(x)
    ```
1. How many are in the `ac_wy` data frame?

```{r}
# to be continued...
```

## 2

1. Use `sessionInfo()` to find out what which packages are currently attached in your R session.
  - How many are there?
  - Run the command `devtools::session_info()`. What's different about the result?
1. Attach the tidyverse package. What does each of the messages mean?:
    ```{r, echo=FALSE}
    library(tidyverse)
    ```
1. How many packages are now attached?
1. Restart your R session and load some **tidyverse** packages individually. Start with **readr**, **dplyr** and **ggplot2** and add more if you need them.
1. Create a barplot showing the number and proportion of crashes in the `ac_wy` dataset on different types of roads using **ggplot2**:
  - Roads with different speed limits (absolute counts and proportions)
  - Different road types (A roads, B roads etc)
    ```{r, echo=FALSE, message=FALSE}
    ac_wy = readr::read_csv("ac_wy.csv")
    ggplot(ac_wy, aes(Speed_limit)) + geom_bar()
    # ggplot(ac_wy, aes(Speed_limit)) +
    #   geom_bar(mapping = aes(y = ..prop.. * 100)) +
    #   ylab("Percentage of roads")
    # ggplot(ac_wy, aes(`1st_Road_Class`)) +
    #   geom_bar() 
    # speed_table = table(ac_wy$`1st_Road_Class`)
    # speed_proportional = speed_table / nrow(ac_wy) * 100
    # speed_table_df = as.data.frame(speed_proportional) 
    # ggplot(speed_table_df, aes(Var1, Freq)) +
    #   geom_bar(stat = "identity")
    ``` 
  


