---
output: github_document
---

<!-- Note: edit the .Rmd file not the .md file -->

## 1.1 R and RStudio

### 1.1a

1. Get into groups of 2 and, in those groups:
1. Create a new RStudio project
1. Create a new blank R script
1. Identify and interact with each of the 4 panels in RStudio
1. Get help on the plot function with `?plot`
1. Create a plot using the `plot()` function
1. Find and install a new package on a topic of your choice with `Tools > Install Packages` (requires internet)
1. Attach the package using `library()`
1. Find and install a new package with `install.packages()`
1. In your source panel write code that creates vector objects `x` and `y` and plots them with `plot(x, y)` to create something that looks like this:
<!-- (is it reproducible?) -->

```{r}
# hint: create a vector object of the numbers 1, 2, 3 and 6 and call it x:
x = c(1, 2, 3, 6)
```

```{r, echo=FALSE}
y = x^2
plot(x, y)
```

- Bonus: find out exactly what R version you are using (tip: use a search engine!)

- Bonus: use R to find out how many minutes you've been alive for. Feel free to use an invented age. Tip: try using both 'base' `as.POSIXct()` and 'tidyverse' `ymd_hm()` functions - you may also need to search online for this.
    
```{r, eval=FALSE, echo=FALSE}
date_of_birth1 = as.POSIXct("1985-12-20 12:00")
date_of_birth2 = ISOdatetime(
  year = 1985,
  month = 12,
  day = 20,
  hour = 12,
  min = 0,
  sec = 0
)
date_of_birth3 = lubridate::ymd_hm("1985-12-20 12:00")
current_time = Sys.time()
diff_days = current_time - date_of_birth1
diff_days * 24 * 60
difftime(current_time, date_of_birth1, units = "s")
```


```{r, echo=FALSE, eval=FALSE}
devtools::install_github("ITSLeeds/stats19")
stats19::dl_stats19_2017_ac()
ac17 = stats19::read_stats19_2005_2014_ac(data_dir = "dftRoadSafetyData_Accidents_2017", filename = "Acc.csv")
ac17 = stats19::format_stats19_2016_ac(ac17)
ac17
View(ac17)
ac_wy = ac17[ac17$Police_Force == "West Yorkshire", ]
ac_wy = dplyr::filter(ac17, Police_)
library(tidyverse)
ac_wy = ac17 %>% 
  filter(Police_Force == "West Yorkshire")
plot(ac_wy$Longitude, ac_wy$Latitude)
```

### 1.1b R classes

1. What type is each of these objects:
    ```{r}
    x = 1:6
    y = sqrt(x)
    z = y + 0.1
    w = c("hello", "hi", 100)
    ```
1. Create a list with the named elements variables `x`, `y` and `z`
1. Import tibble with `library(tibble)` and swap `list` for `tibble`
1. Fetch `ac_wy.csv` from our releases page: "https://github.com/ITSLeeds/highways-course/releases/download/0.2/ac_wy.csv"
1. You can download it with R like this:
    ```{r, eval=FALSE}
    f = "https://github.com/ITSLeeds/highways-course/releases/download/0.2/ac_wy.csv"
    download.file(url = f, destfile = "ac_wy.csv")
    ```
1. Read the csv with read_csv()
1. Find the number of rows in the tibble (search the internet to find the right function)
1. Work out what the summary function does when applied to the tibble and on columns of the tibble
1. Find all the fatal accidents by subsetting with `[` or using the `filter` function from dplyr
1. Create summary stats for all accidents involving more than 2 pedestrians

```{r, echo=FALSE}
# to be continued...
```

## 2 Stats refresher and packages

1. Discuss in groups: what kind of statistical analysis do you you do, exploratory or hypothesis testing?
1. Use `sessionInfo()` to find out what which packages are currently attached in your R session.
    - How many are there?
    - Run the command `devtools::session_info()`. What's different about the result?
1. Attach the tidyverse package. What does each of the messages mean?:
    ```{r, echo=FALSE}
    library(tidyverse)
    ```
1. How many packages are now attached?
1. Restart your R session and load some **tidyverse** packages individually. Start with **readr**, **dplyr** and **ggplot2**.
1.  Run `ggplot(data = mpg)`. What do you see?
1.  What does the `drv` variable describe?  Read the help for `?mpg` to find
    out.
1.  Make a scatterplot of `hwy` vs `cyl`.
1. Create a barplot showing the number and proportion of crashes in the `ac_wy` dataset on different types of roads using **ggplot2**:
  - Roads with different speed limits (absolute counts and proportions)
  - Different road types (A roads, B roads etc)
    ```{r, echo=FALSE, message=FALSE}
    ac_wy = readr::read_csv("ac_wy.csv")
    ggplot(ac_wy, aes(Speed_limit)) + geom_bar()
    ggplot(ac_wy, aes(Speed_limit)) +
      geom_bar(mapping = aes(y = ..prop.. * 100)) +
      ylab("Percentage of roads")
    ggplot(ac_wy, aes(`1st_Road_Class`)) +
      geom_bar()
    speed_table = table(ac_wy$`1st_Road_Class`)
    speed_proportional = speed_table / nrow(ac_wy) * 100
    speed_table_df = as.data.frame(speed_proportional)
    ggplot(speed_table_df, aes(Var1, Freq, fill = "orange")) +
      geom_bar(stat = "identity") +
      ylab("Percent") +
      xlab(NULL) +
      theme(axis.text.x = element_text(angle = 90))
    ``` 
  
# 3 Spatial data

1. Practical: [Section 3.2 to 3.2.2](https://geocompr.robinlovelace.net/attr.html#vector-attribute-manipulation) of handouts

1. Work through exercises 1:3 in the hand-outs
1. How many states:
    - Contain the letter E (hint: `?grepl`)?
    - Start with A (hint: search for "regex starts with")?
1. Plot all the 'E' states in red and plot the 'A' states with a thick border.
    ```{r, message=FALSE, echo=FALSE}
    library(sf)
    library(spData)
    library(dplyr)
    us_e = us_states %>% 
      filter(grepl(pattern = "e", x = NAME, ignore.case = T))
    us_a = us_states %>% 
      filter(grepl(pattern = "^A", x = NAME))
    plot(us_states$geometry)
    plot(us_e$geometry, col = "red", add = TRUE)
    plot(us_a$geometry, lwd = 5, add = TRUE)
    ```
- Bonus: use the `world` dataset to find the top 3 smallest and largest countries in the world by population and area, and plot them on a single map.

## 4 Spatial data II and roadworks

1. Practical: working-through sections [3.2.3 to 3.2.4](https://geocompr.robinlovelace.net/attr.html#vector-attribute-joining) of hand-outs
1. [Exercises](https://geocompr.robinlovelace.net/attr.html#exercises-1): 4 to 6 onwards
1. Identify the states that grew most and least from 2010 to 2015 and plot them.
    ```{r, message=FALSE, echo=FALSE}
    library(tmap)
    us_states$change = us_states$total_pop_15 - us_states$total_pop_10
    us_states$growth = NA
    top3 = tail(order(us_states$change), 3)
    bot3 = head(order(us_states$change), 3)
    us_states$growth[top3] = "Most"
    us_states$growth[bot3] = "Least"
    tm_shape(us_states) +
      tm_polygons("growth")
    ```
1. Create an obect that represents the boundary of the USA.
- Bonus: create a random distribution of 1000 points across the United States and subset those that are in Texas. Plot the results to check your code works, does it look something like this?
    ```{r, message=FALSE, echo=FALSE}
    us_union = st_union(us_states)
    xran = runif(n = 1000, min = st_bbox(us_states)[1], max = st_bbox(us_states)[3])
    yran = runif(n = 1000, min = st_bbox(us_states)[2], max = st_bbox(us_states)[4])
    mran_sample = st_sample(us_states, 1000)
    mran = data.frame(xran, yran, n = 1:1000)
    crs_us = st_crs(us_states)
    mran_sf = st_as_sf(mran, coords = c("xran", "yran"), crs = crs_us)
    # st_crs(mran_sample)
    mran_sf = st_sf(geometry = mran_sample)
    tex = us_states %>% 
      filter(NAME == "Texas")
    sel_tex = st_within(mran_sf, tex)
    sel_tex_binary = lengths(sel_tex) > 0
    mran_sf$in_texas = sel_tex_binary
    tm_shape(us_union) +
      tm_polygons() +
      tm_shape(mran_sf) +
      tm_dots(col = "in_texas", size = 0.2)
    ```
- Advanced option 1: Section [4.2 - Spatial operations on vector data](https://geocompr.robinlovelace.net/spatial-operations.html#spatial-vec) of Geocomputation with R
- Advanced option 2: Install the **roadworksUK** package and identify which MSOA in Ashford had the highest number of gas-related roadworks in the `htdd_ashford` dataset.
