---
output: github_document
---

<!-- Note: edit the .Rmd file not the .md file -->

## 5 Point data

1. Starting where we left-off, complete exercises 7 and 12 from the hand-outs
3. Read-in the dataset `leeds.Rds`
1. Create a new column in the dataset called `area` that reports the area of each MSOA, and plot the result (hint: use `st_area()`)
    ```{r}
    library(sf)
    leeds = readRDS("data/leeds.Rds")
    leeds$area = st_area(leeds)
    plot(leeds["area"])
    ```

2. Create an object called `ac_wy_sf`, a spatial version of the `ac_wy` dataset read-in previously (hint: use the function `st_as_sf()` and coordinate variables Location_Easting_OSGR/Location_Northing_OSGR --- don't forget to set the CRS to 27700)
3. Plot the result, showing only the accident severity.
    ```{r, echo=FALSE}
    ac_wy = readRDS("ac_wy.rds")
    ac_wy_sf = st_as_sf(ac_wy,
                        coords = c("Location_Easting_OSGR", "Location_Northing_OSGR"),
                        crs = 27700)
    plot(ac_wy_sf["Accident_Severity"])
    ```
4. Create a new object called `ac_wy_wgs` that tranforms the `ac_wy_sf` into a geographic (lat/lon) CRS (hint: use `st_crs()` and the EPSG code 4326)
5. Find out how many fatal crashes happened in each of the MSOA areas in Leeds
    ```{r, echo=FALSE}
    # library(sf)
    # ac_wy = readRDS("ac_wy.rds")
    # ac_wy_sf = st_as_sf(ac_wy,
    #                     coords = c("Location_Easting_OSGR", "Location_Northing_OSGR"),
    #                     crs = 27700)
    # plot(ac_wy_sf["Accident_Severity"])
    ```