---
output: github_document
---

<!-- Note: edit the .Rmd file not the .md file -->

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
```

# Exercises day 2 - Transport

```{r, include=FALSE}
# vignette(package = "stplanr")
library(sf)
library(stplanr)
routes_fast_sf$lengths = st_length(routes_fast_sf) %>% 
    as.numeric()
rf = filter(routes_fast_sf, lengths > 0)
rf_top3 = rf %>% 
    top_n(n = 12, wt = lengths)
rf_bot3 = rf %>% 
    top_n(n = 12, wt = desc(lengths))

```

- Load the **stplanr** library
- How many datasets does it contain (hint use the command `data(package = ...)`? Take a look at the help for a few of them.
- Type in `stplanr::` in your console and hit TAB. What do you see? By scrolling down find functions beginning with `geo_*`
- Find the vignettes in the **stplanr** library and open the vignette called `introducing-stplanr`.
- Using the function `st_length()` find the length of each route in `routes_fast_sf` and save the result in an object called `routes_fast_sf$lengths`
- Find the `summary()` of the `rf_lengths`
- How many lines have length = 0?
- Create an object called `rf` that only contains routes with non-zero distances
- Identify the top 3 longest and top 3 shortest routes and plot them over the routes fast. The result should look like something like this:

```{r}
plot(rf$geometry, col = "grey")
plot(rf_top3$geometry, col = "red", add = TRUE)
plot(rf_bot3$geometry, col = "blue", add = TRUE)
```

- Create a single desire line for the 2nd to 9th flow using **stplanr** code:

- Convert the result to `sf` class with `st_as_sf()` (**stplanr** solution for the 12th line below - the result should look something like that below):

```{r, eval=FALSE}
od2line(flow = flow[12,], zones = cents)
```

```{r}
l = od2line(flow[2:9, ], cents)
l_sf = st_as_sf(l)
plot(l_sf)
```



- **Advanced**: Create a single desire line for the 12th flow **without** using **stplanr** code:

- Visualise the `flowlines` object using **tmap** with different colours and widths
- Create a 'oneway' version of the `flowlines` and visualise these (tricky)
- Create a **sf** version of the `flowlines` object (advanced)

- Work through the [vignette](https://cran.r-project.org/web/packages/stplanr/vignettes/introducing-stplanr.html) provided by:

```{r, eval=FALSE}
vignette("introducing-stplanr")
```

- For another day (or if you're very fast!): read-over https://geocompr.robinlovelace.net/transport.html and complete the exercises at the end