---
output: github_document
# output: word_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=FALSE, echo=FALSE}
# generate/publish online the materials
# piggyback::pb_new_release(tag = "0.1")
file.rename("README.docx", "highways-course.docx")
piggyback::pb_upload("highways-course.docx")
knitr::purl("README.Rmd")

# to concatenate files (in bash)
cat slides/intro.Rmd slides/stats-pkgs.Rmd README.Rmd > /tmp/all.Rmd

# save html slides
wd_old = setwd("slides")
f = list.files(pattern = "spatial")
zip(zipfile = "spatial.zip", files = f)
f = list.files(pattern = "roadworks")
zip(zipfile = "roadworks.zip", files = f)
setwd(wd_old)

mv -v slides/*.zip . # in bash

piggyback::pb_upload("roadworks.zip")
piggyback::pb_upload("spatial.zip")

citr::tidy_bib_file(rmd_file = "/tmp/all.Rmd", messy_bibliography = "~/allrefs.bib", file = "slides/refs.bib")

# save leeds data
library(sf)
leeds_bound = ukboundaries::leeds
leeds = ukboundaries::msoa2011_lds
st_crs(leeds)
leeds_cents = st_centroid(ukboundaries::msoa2011_vsimple)[leeds, ]
plot(leeds_cents)
leeds = ukboundaries::msoa2011_vsimple[ukboundaries::msoa2011_vsimple$objectid %in% leeds_cents$objectid, ]
plot(leeds)
saveRDS(leeds, "data/leeds.Rds")
file.size("data/leeds.Rds") / 1e6
```

## Aim

The aim of R for Highways Research and Transport Planning is to teach how R can be used as a powerful tool for processing, visualising and modelling transport data.

Course contents can be found online at: https://github.com/ITSLeeds/highways-course

## Location

Leeds Institute for Transport Studies, 40 University Road, LS2 9JT.
It can be seen in this [web map](https://www.openstreetmap.org/way/84749920), highlighted in red below, just opposite Woodhouse Moore (it's a 20 minute walk from the station):

```{r itsmap, include=FALSE}
library(osmdata)
library(tmap)
its = opq("Leeds") %>% 
  add_osm_feature(key = "name", value = "Institute for Transport Studies") %>% 
  osmdata_sf()
# m = mapview::mapview(its$osm_polygons, zoom = 16)
m = tm_shape(its$osm_polygons) +
  tm_polygons(col = "red") +
  tm_view(set.view = 16)
tmap_mode("view")
ml = tmap_leaflet(m)
```

```{r, echo=FALSE}
ml
```

## Course contents

09:00-09:30 Arrival and set-up

09:30-11:00 Introduction to the course and software

- Introduction to R
- R installation questions/debugging
- How to use RStudio (practical in groups of 2)
- R classes and working with data frames (CC)

11:15-12:30 Statistics and packages

<!-- And example from the PCT -->
- Stats refresher: plots and descriptive statistics
- Predictive models
- Using packages: examples with the tidyverse

**Lunch**

13:30-15:00 Spatial data in R

- Spatial data in R
- R's spatial ecosystem (see section [1.4 of Geocomputation with R - package ecosystem](https://geocompr.robinlovelace.net/intro.html#rs-spatial-ecosystem))
- Practical: [Section 3.2 to 3.2.2](https://geocompr.robinlovelace.net/attr.html#vector-attribute-manipulation) of handouts
- [Exercises](https://geocompr.robinlovelace.net/attr.html#exercises-1): 1 to 3 in the hand-outs
- Further reading: [Chapter 7](https://geocompr.robinlovelace.net/read-write.html) of Geocomputation with R - data import/export and [exercises](https://geocompr.robinlovelace.net/read-write.html)
- Bonus (relies on internet/data): read-in and analyse datasets you use at work as data frames or sf objects. 
  
15:15-16:30 Real-work example and consolidation

- Demo of roadworks data with R
- Practical: working-through sections [3.2.3 to 3.2.4](https://geocompr.robinlovelace.net/attr.html#vector-attribute-joining) of hand-outs
- [Exercises](https://geocompr.robinlovelace.net/attr.html#exercises-1): 4 to 6 onwards
- Advanced option: Section [4.2 - Spatial operations on vector data](https://geocompr.robinlovelace.net/spatial-operations.html#spatial-vec) of Geocomputation with R
<!-- - Advanced option: Section [2.2](https://geocompr.robinlovelace.net/spatial-class.html#vector-data) of Geocomputation with R -->

**Day 2 transport data**

09:30-11:00 Stream 1: An introduction to point (Stats19) data

- Point data: Stats19
- Spatial and temporal subsetting
- Aggregation

09:30-11:00 Stream 2: Recap of day

Slow demo of:

- Set-up
- Subsetting
- Dplyr and ggplot2

11:15-12:30 Desire lines and routing

- Desire lines: using origin-destination data
- Routing
- Bonus: finding crash hotspots

**Lunch**

13:30-15:00 Road traffic data

- Traffic data introduction (Josh Manning and Ivo)
- Temporal analysis
- Spatial analysis

15:15-16:30 Practical application

- Stream 1: Consolidation of exercises
- Stream 2: Working on real datasets such as flooding or traffic data, or improving specific skills such as visualisation

## Prerequisites

### Prior reading/experience

If you are new to R, ensure you have attempted a basic introductory course such as DataCamp's [introduction to R](https://www.datacamp.com/courses/free-introduction-to-r) free course or equivalent.

If you're interested in R for 'data science' and installing/updating/choosing R packages, these additional resources are recommended (these optional resources are all freely availble online):

- The introductory chapter of [R for Data Science](https://r4ds.had.co.nz/introduction.html)
- Chapter 2 on [setting-up R](https://csgillespie.github.io/efficientR/set-up.html) and section 4.4 on [package selection](https://csgillespie.github.io/efficientR/workflow.html#package-selection) in the book *Efficient R Programming*
- The [Transport](https://geocompr.robinlovelace.net/transport.html) chapter of the new book Geocomputation with R

<!-- (MIDAS Gold) -->

<!-- ## Optional extras (to discuss) -->

<!-- - Roadworks data (HTDD/Scottish/Leeds data - HE have data?) -->
<!-- - Stats19 -->
<!-- - Routing engines -->
<!-- - Air pollution -->
<!-- - Traffic data (other) -->

<!-- ## To discuss/confirm -->

<!-- - 10 ppl HE + 8 RAC  -->
<!-- - Managed work laptops - install pre-requisites - pre-reqs document. -->
<!-- - Demonstrators (ask Josh - possible fee, ask Maxine should be fine) -->
<!-- - Ivo Helper -->
<!-- - Location: look into it - plus refreshments -->
<!--   - None HE Leeds -->
<!--   - Maybe HE Birmingham -->
<!--   - Maybe RAC -->
<!-- - Timing: mid November or w/c 10th Dec -->

### Computing requirements

Attendees are expected to bring their own laptop with the following packages installed and working.
You can check these are all installed, and install those that are not installed, as follows (you can also just type `install.packages("sf")` etc):

```{r packages, message=FALSE, warning=FALSE, eval=FALSE}
install.packages("osmdata")   # for working with open street map data
install.packages("sf")        # a package for working with spatial data
install.packages("spData")    # provides example data
install.packages("stplanr")   # a transport data package
install.packages("tidyverse") # metapackage for data science
install.packages("tmap")      # a mapping package

# Make sure your packages are up-to-date with:
update.packages()
```



In addition, it would be useful to have oneminutetraffic, which can be installed with:

```{r, message=FALSE}
devtools::install_github("RACFoundation/oneminutetrafficdata")
```

## Reproducible example

The code in the following example checks you have the necessary packages installed.
It results in a map that will guide you to the location of the course.

Attach the packages:

```{r, message=FALSE}
library(sf)
library(stplanr)
library(tidyverse)
```

The overall route assuming you're travelling from London:

```{r}
uk = spData::world %>% 
  filter(name_long == "United Kingdom")
origin_lnd = c(-0.1, 51.5)
destination = c(-1.55, 53.8)
odmatrix = matrix(c(origin_lnd, destination), ncol = 2, byrow = TRUE)
line_lnd = st_linestring(odmatrix) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
plot(st_geometry(uk), col = "grey")
plot(line_lnd, add = TRUE, lwd = 5, col = "red")
```


```{r, echo=FALSE, eval=FALSE}
origin_lds = geo_code("Leeds rail station")
destination = geo_code("Worsley Building, Leeds")
bb = matrix(c(-1.56, 53.7, -1.53, 53.9), ncol = 2)
m1 = tmap::qtm(line_lnd)
tmap::tmap_leaflet(m1)

library(osmdata)
roads = opq(bbox = bb) %>% 
  add_osm_feature(key = "highway", value = "pri|sec|res", value_exact = FALSE) %>% 
  osmdata_sf()
```


```{r, echo=FALSE, eval=FALSE}
sln = SpatialLinesNetwork(roads$osm_lines)

from_sln = find_network_nodes(sln, origin_lds[1], origin_lds[2])
to_sln = find_network_nodes(sln, destination[1], destination[2])

r_local = sum_network_routes(sln, from_sln, to_sln, "length", combinations = F)

m2 = tmap::qtm(r_local)
library(leaflet)
tmap::tmap_leaflet(m2) %>% 
  addCircleMarkers(lng = origin_lds[1], lat = origin_lds[2]) %>% 
  addCircleMarkers(lng = destination[1], lat = destination[2])
```

```{r, eval=FALSE, echo=FALSE}
# bumpf (not used)
getbb_of_two_points = function(p1, p2) {
  minx = min(c(p1[1], p2[1]))
  maxx = max(c(p1[1], p2[1]))
  miny = min(c(p1[2], p2[2]))
  maxy = max(c(p1[2], p2[2]))
  rbind(
    c(minx, maxx),
    c(miny, maxy)
  )
}
getbb_of_two_points(origin_lds, destination)
mat_orig = matrix(origin_lds, ncol = 2)
origin_sf = st_point(origin_lds) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
destination_sf = st_point(destination) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
mat_dest = sln@sl
nearest_o = knn_orig <- nabor::knn(mat_via, query = mat_orig, k = 1)$nn.idx
route_lds = (from = origin_lds, to = destination)
```

## Shortcut keys:

```{r}
# Useful shortkeys --------------------------------------------------------

# Tab-Autocomplete
# Alt-Up/Down: move current line (shift copies it)
# Ctl-Shift-C: Comment line
# Ctl-Shift-R: Add section
# Ctl-Shift-m %>% %>% %>% %>% %>% %>% %>% %>% %>% 
# Ct1-1/2 put cursor in source/console
# Ctl-Alt-T Execute as system command, e.g.:
# cp -Rv * /tmp
# Ctl-Alt-Down: duplicate cursor
# Alt-Shift-K: show autocomplete keys!
```


<!-- Note: you can test all of these things work by running the following command: -->

<!-- ```{r, eval=FALSE} -->
<!-- source("https://raw.githubusercontent.com/ITSLeeds/highways-course/master/README.R") -->
<!-- ``` -->

<!-- Bonus: find the route from Leeds rail station  -->
<!-- (see the code in `README.R` on the course website at https://github.com/ITSLeeds/highways-course ) -->
