---
output: github_document
# output: word_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Aim

The aim of highways-course is to provide content for a 2 day course on R and transport data with R.

## Course contents (times TBC)

- Introduction to R
- How to use RStudio 
- Using packages (emphasis: tidyverse)
- Stats refresher
- Temporal data

**Lunch**

- Spatial data in R
- Intro + R's spatial ecosystem: 
       See section 1.4 of Geocomputation with R
     Working with attribute data (10:30 - 11:00)
         Section 3.2 of handouts

- Spatial data and the tidyverse lecture (11:15 - 11:30)
  - Vector data classes
  - Online tutorial: 2.1 of Geocomputation with R (11:30 - 12:00)

**Day 2 transport data**

- Origin-destination / stats19

**Lunch**

- Traffic data (MIDAS Gold)

## Optional extras (to discuss)

- Roadworks data (HTDD/Scottish/Leeds data - HE have data?)
- Stats19
- Routing engines
- Air pollution
- Traffic data (other)

## To discuss/confirm

- 10 ppl HE + 8 RAC 
- Managed work laptops - install pre-requisites - pre-reqs document.
- Demonstrators (ask Josh - possible fee, ask Maxine should be fine)
- Ivo Helper
- Location: look into it - plus refreshments
  - None HE Leeds
  - Maybe HE Birmingham
  - Maybe RAC
- Timing: mid November or w/c 10th Dec

## Prerequisites

Attendees are expected to bring their own laptop with the following packages installed and working:

```{r}
pkgs = c(
  "osmdata",   # for working with osmdata
  "sf",        # a package for working with spatial data
  "stplanr",   # a transport data package
  "tidyverse", # metapackage for data science
  "tmap"       # a mapping package
)
```

You can check these are all installed, and install those that are not installed, as follows (you can also just type `install.packages("sf")` etc):

```{r}
pkgs_installed = pkgs %in% installed.packages()
names(pkgs_installed) = pkgs
pkgs_installed
if(!all(pkgs_installed)) {
  install.packages(pkgs[!pkgs_installed])
}
```

In addition, it would be useful to have oneminutetraffic, which can be installed with:

```{r}
devtools::install_github("RACFoundation/oneminutetrafficdata")
```


## Reproducible example

The code in the following example checks you have the necessary packages installed.
It results in a map that will guide you to the location of the course.

Attach the packages:

```{r}
library(sf)
library(stplanr)
library(tidyverse)
```

The overall route assuming you're travelling from London:

```{r}
origin_lnd = geo_code("London Kings Cross")
destination = geo_code("Worsley Building, Leeds")
odmatrix = matrix(c(origin_lnd, destination), ncol = 2, byrow = TRUE)
line_lnd = st_linestring(odmatrix) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
m1 = tmap::qtm(line_lnd)
tmap::tmap_leaflet(m1)
```

The route from Leeds rail station looks like this:

```{r}
origin_lds = geo_code("Leeds rail station")
destination = geo_code("Worsley Building, Leeds")
bb = matrix(c(-1.56, 53.7, -1.53, 53.9), ncol = 2)
library(osmdata)
roads = opq(bbox = bb) %>% 
  add_osm_feature(key = "highway", value = "pri|sec|res", value_exact = FALSE) %>% 
  osmdata_sf()
sln = SpatialLinesNetwork(roads$osm_lines)

from_sln = find_network_nodes(sln, origin_lds[1], origin_lds[2])
to_sln = find_network_nodes(sln, destination[1], destination[2])

r_local = sum_network_routes(sln, from_sln, to_sln, "length", combinations = F)

m2 = tmap::qtm(r_local)
library(leaflet)
tmap::tmap_leaflet(m2) %>% 
  addCircleMarkers(lng = origin_lds[1], lat = origin_lds[2]) %>% 
  addCircleMarkers(lng = destination[1], lat = destination[2])
```




```{r, eval=FALSE, echo=FALSE}
getbb_of_two_points = function(p1, p2) {
  minx = min(c(p1[1], p2[1]))
  maxx = max(c(p1[1], p2[1]))
  miny = min(c(p1[2], p2[2]))
  maxy = max(c(p1[2], p2[2]))
  rbind(
    c(minx, maxx),
    c(miny, maxy)
  )
}
getbb_of_two_points(origin_lds, destination)
mat_orig = matrix(origin_lds, ncol = 2)
origin_sf = st_point(origin_lds) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
destination_sf = st_point(destination) %>% 
  st_sfc() %>% 
  st_sf(crs = 4326)
mat_dest = sln@sl
nearest_o = knn_orig <- nabor::knn(mat_via, query = mat_orig, k = 1)$nn.idx
route_lds = (from = origin_lds, to = destination)
```



