---
title: 'R for Highways Research and Transport Planning'
subtitle: "R datatypes"
author: "Colin Caine, School of Geography"
date: "University of Leeds, `r Sys.Date()`."
output: ioslides_presentation
widescreen: true
logo: its-logo-square.png
---

```{r, echo=FALSE, eval=FALSE}
file.copy("slides/R-datatypes.html", ".", T)
piggyback::pb_upload("R-datatypes.html")
piggyback::pb_upload("2017-12-01--2017-12-07.7z")
```

## Learning objectives

- Know and understand the basic R types
  - Appreciate how they correspond to formats you are already familiar with (e.g. csv)
- Understand vectorisation
- Be able to filter and select from R types


## What types are these?

```{r}
x = 1
x
y = c(1, 2, 3, 9, 8, 7)
y
```

## What types are these?

```{r}
x = 1
typeof(x)
y = c(1, 2, 3, 9, 8, 7)
typeof(y)
```

What's funny about this?

<!--
What's a double?
  Double precision float IEEE 754.
Why not integer?
  R assumes numeric data is
Why the same for scalar and vector?
-->

## What's a double?

> - A way of encoding a wide range of numbers (0.025, 10 million and 1, 1.4032e36)
> - Fraction * exponent (0.25 * 10e-1, 1 * 10e7)
> - The fraction may be an approximation
> - Double precision floating point number (IEEE 754)

## Why not an integer?

- A double is generally what you want

> - Just don't worry about it

## A vector programming language

```{r}
# Vector creation shorthand
1:10
```

```{r, eval=F}
# Vector-scalar multiplication
1:10 * 2
# Vector-vector multiplication
1:3 * 1:3
```

## A vector programming language

```{r}
# Vector scalar multiplication
1:10 * 2
# Vector-vector multiplication
1:3 * 1:3
```

- It's **not** the dot or cross product

## A vector programming language

```{r}
# ???
1:10 * 1:5
```

> - "broadcasting"
> - "recycling"

## *Why* a vector programming language?

- *Element-wise* operation is handily expressive
- End up operating on whole columns or tables at once
- Usually what you want in stats

```{r}
# Center a distribution on 0 and compress it's spread
x = 0:10
(x - mean(x)) / 2
```


## The common atomic vector types

```{r}
typeof(c("a", "b", "sea"))
typeof(c(TRUE, FALSE, TRUE))
typeof(1:10)
typeof(c(1, 2, 3))
```


## Why atomic?

- They can only contain values of a single type


## What other data types might we need?

- Can you represent the data you use at work as vectors?

> - What about GPS traces?

## Lists

- Collect many vectors together (e.g. GPS traces)
- Or a list of values of different data types (e.g. a csv)

```{r}
list(c(1, 2, 3, 4), c("one", "two", "three", "four"))

users = 1:100
gps_traces = list(list(x = c(1, 1, 1, 2, 2, 1, 1), y = c(5, 4, 5, 5, 6, 7, 6)))
```

## Extracting items

```{r}
x = 1:3 * 33
x[1] # subset vector by index
x[c(1,2)] # subset by vector
```

## Extracting items

```{r}
x > 45
x[x > 45] # subset by logical vector
```

## Extracting items

With lists we sometimes use `[[` instead

`[` subsets the object so we end up with the same type of object back

```{r}
y = list(1:3, c(1, 10, 50))
y[1]
y[[1]]
```

## Extracting items

```{r}
y[1][1]
y[[1]][1]
y[[1]][[1]]
```

## Extracting items

Index by name. Is this like `[` or `[[`?

```{r}
y = list(ids = 1:3, price = c(1, 10, 50))
y$ids
y$price
```


## Data frames

<div style="float: left; width: 50%;">

 - Most famous of R's types
 - Represents a table of data e.g:

```{r, echo=FALSE}
library(ggplot2)
as.data.frame(unique(mpg[1:2])[1:10,])
```

</div>

<div style="float: left; width: 50%;">

> - What does this look most like?
> - It's mostly just a list of vectors
> - But all the vectors must be the same length
> - You'll see a lot of these later

</div>

## Tibbles

- Generally prefer tibbles

```{r, echo=FALSE}
mpg
```

## Get data from dataframes and tibbles

```{r, eval=FALSE}
library(ggplot2) # provides the mpg dataset
mpg[c(1,2)] # Select columns 1 and 2
mpg[3:5] # Columns 3, 5
mpg[1:10, 1:2] # Columns 1 and 2 of rows 1-10
```


## Review

> - Four common vector types
>   - integer, double, logical, character
> - Operations are element-wise
> - Use tibbles to represent data tables
> - Use lists when you want to mix objects of different types
>   - Or collect collections of different lengths
> - subset collections with `[`
> - extract items from collections with `[[` and `$name`


## Gotchas

```{r}
TRUE == T
FALSE == F
c(1, 2, "three")
1 == "1"
```


<!--
Deliberately not covered due to time constraints:

- Matrices
- Functions/closures
- OOP stuff
-->
